\count0=1
\magnification=\magstep1
\baselineskip=14pt
\overfullrule=0pt
\rightline{Revised September 2023}
\centerline{\bf DAVID EISENBUD}
\centerline{eisenbud.io.github.com}


\centerline{SELECTED PUBLICATIONS}
\vskip .5cm
\centerline{\bf{10 Representative Publications}}
\vskip .3cm
\newcount\pno
\pno=0

\advance\pno by 1
\item{\the\pno.}
What Makes a Complex Exact?  
(with D.~A.~Buchsbaum), J.~of Alg., 25 (1973) 
259--268.

\advance\pno by 1
\item{\the\pno.}
Every Algebraic Set in $n$-Space 
is the Intersection of $n$ Hypersurfaces 
(with E.~G. Evans), Invent.~Math.~19 (1973) 107--112.

\advance\pno by 1
\item{\the\pno.}
An algebraic formula for the degree of a $C^{\infty}$
 map germ (with H.~Levine), 
Annals of Math.~106 (1977) 19--44.

\advance\pno by 1
\item{\the\pno.}
Homological Algebra over a Complete Intersection, 
Trans.~Am.~Math.~Soc.~260 
(1980), 35--63.

\advance\pno by 1
\item{\the\pno.}
  Gorenstein ideals of height 3 (with D.A. Buchsbaum), 
  In Seminar D. Eisenbud/B. Singh/W. Vogel, Vol. 2, volume 48 or Teubner-Texte zur Math., (30--48.  Teubner, Leipzig, (1982).

\item{\the\pno.} 
{\bf Three dimensional Link Theory 
and Invariants of Plane Curve Singularities}, 
(with W.~Neumann), Annals of Math.~Studies 110, Princeton 
University Press Princeton NJ (1985).

\advance\pno by 1
\item{\the\pno.}
The Kodaira dimension of the 
moduli space of curves of genus  $\geq 23$ 
(with J.~Harris),  Invent.~Math.~90 (1987) 359--387.

%\advance\pno by 1
%\item{\the\pno.}
%Open Problems in Computational Algebraic 
%Geometry and commutative 
%Algebra, in 
%{\bf Computational Algebraic Geometry and Commutative Algebra,
%Cortona 1991},
%(ed.~D.~Eisenbud and L.~Robbiano) 
%Cambridge University Press, Cambridge, England, (1993) 49--71.

\advance\pno by 1
\item{\the\pno.}
{\bf Commutative Algebra With A 
View Toward Algebraic Geometry},  
Graduate Text 150, 
Springer-Verlag, 1995.

\advance\pno by 1
\item{\the\pno.}
Exterior algebra methods for the minimal resolution conjecture (with
Sorin Popescu, Frank-Olaf Schreyer, and Charles Walter).  Duke
Math. J.  112 (2002), no. 2, 379--395.

\advance\pno by 1
\item{\the\pno.}  
Resultants and Chow forms via exterior syzygies. 
(with F.-O. Schreyer\ and\ J. Weyman).
J. Amer. Math. Soc. {\bf 16} (2003), 
537--579.

 \advance\pno by 1
\item{\the\pno.}
 Betti Numbers of Graded Modules and Cohomology of Vector Bundles
 (with Frank-Olaf Schreyer).
 Jour. Amer. Math. Soc., (2009),  no. 3, 859--888.

\vfill\eject

\goodbreak
\centerline{\bf{Publications, Last Three Years}}
\vskip .3cm
\newcount\pno
\pno=1

\item{\the\pno.}
Quadratic complete intersections. (with I. Peeva and F.-O. Schreyer) J. Algebra 571 (2021), 15--31.


\advance\pno by 1
\item{\the\pno.}
Layered resolutions of Cohen-Macaulay modules (with I. Peeva). J. Eur. Math. Soc. (JEMS) 23 (2021), no. 3, 845--867.

\advance\pno by 1
\item{\the\pno.}
Tate resolutions and MCM approximations. Commutative algebra—150 years with Roger and Sylvia Wiegand, 35--47, Contemp. Math., 773,

\advance\pno by 1
\item{\the\pno.}
 Classical Algebraic Geometry (O. Debarre, G. Farkas and R. Vakil; Oberwolfach Rep. 18 (2021), no. 2, 1519--1577. 

\advance\pno by 1
\item{\the\pno.}
Remembering David Buchsbaum (with J. Weyman) Notices Amer. Math. Soc. 69 (2022), no. 1, 76--87.

\advance\pno by 1
\item{\the\pno.}
Mathematics at the airport? Why sure! (with N. Mullen and C. Stoll)
Notices Amer. Math. Soc. 69 (2022), no. 4, 612--615.

\advance\pno by 1
\item{\the\pno.}
Linearity of free resolutions of monomial ideals (with H. Dao) Res. Math. Sci. 9 (2022), no. 2, Paper No. 35, 15 pp. 

\item{\the\pno.}
Residual Intersections of 2xn determinantal ideals  (with Bernd Ulrich) 
submitted to the volume in honor of Shiing-Shen Chern, accepted for the Journal Algebraic Geometry and Physics, Tsinghua, China.

\advance\pno by 1
\item{\the\pno.}
Burch Index, Summands of Syzygies and Linearity in Resolutions (with H. Dao) submitted to Bulletin of the Iranian Mathematical Society.

\advance\pno by 1
\item{\the\pno.}
F.S. Macaulay: From Plane Curves to Gorenstein Rings (with Jeremy Gray) Bull. Am. Math. Soc..

\advance\pno by 1
\item{\the\pno.}
Residual Intersections and Linear Powers (with C. Huneke, B. Ulrich) To appear, Trans. Am. Math. Soc.

\advance\pno by 1
\item{\the\pno.}
{\bf Practical Curves} (Graduate Textbook, approximately 400 pages) To appear, American Mathematical Society.
\end
